---
title: "Metro Van - rent analysis"
output: html_notebook
---

Author: Martin Monkman
Date: 2016-12-06

---

This file reads rent data (advertisements) from the Metro Vancouver region.




```{r}

library(dplyr)
library(readr)

columnnames <- c("reg", "latitude", "longitude", "rent", "bedrooms", "bathrooms", "sq_ft")

# read files, add column names
data1 <- read_csv("cl_2016_4_3.csv", col_names = FALSE)
colnames(data1) <- columnnames

data2 <- read_csv("cl_2016_4_4.csv", col_names = FALSE)
colnames(data2) <- columnnames


# bind the two data frames
# https://rdrr.io/cran/dplyr/man/bind.html
f01_data <- unique(bind_rows(data1, data2))
f01_data$rent_num <- gsub('\\$','', f01_data$rent)


# filter out records that are not complete (i.e. NA values)
f02_data <- f01_data[complete.cases(f01_data), ]


# filter so that bedrooms = 1 or 2, bathrooms = 1
f03_data <- f02_data %>%
  filter(bedrooms != 0)

f03_data_1.2br <- f03_data %>%
  filter(bedrooms <= 2 & bathrooms == 1)


```

## Spatial data (mapping)


### Read in the shape file

For this exercise, we'll use the census tract shape files for the 2011 Census of Canada.


**References**

http://www.studytrails.com/r/r-and-gis-working-with-shapefiles/



```{r}
#
library(maptools)

library(sp)
library(rgeos)
#library(rgdal)

# from http://www.studytrails.com/r/r-and-gis-working-with-shapefiles/

crswgs84=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# read shapefile
BC_censustracts <- readShapePoly("CTshape/CT_2011.shp",proj4string=crswgs84,verbose=TRUE) 
 
# examine shapefile
class(BC_censustracts)
str(BC_censustracts@data)

# plot shapes
plot(BC_censustracts)


```


Zoom in on Metro Vancouver

Simple: just define the x and y limits using longitude (as x) and latitude (as y)

```{r}

# plot shapes
plot(BC_censustracts, 
     xlim=c(-123.3,-122.7), ylim=c(48.97,49.4))

```


### Spatial join

The next step is to use the census tract shape file to tell us the census tract the rental unit (in the data file f03_data_1.2br).

This requires a spatial join, using the `over` command in the package sp.



---

#### futher reading

http://benrifkind.github.io/Mapping-Stats-Canada-Data-with-R-Part-3-of-3/


https://csde.washington.edu/services/gis/workshops/SPATIALR.shtml 

http://gis.stackexchange.com/questions/19064/how-to-open-a-shapefile-in-r


http://gis.stackexchange.com/questions/137621/join-spatial-point-data-to-polygons-in-r

https://cran.r-project.org/web/packages/sp/index.html 

http://stackoverflow.com/questions/24174042/matching-georeferenced-data-with-shape-file-in-r


**Statistics Canada shapefiles**

https://www12.statcan.gc.ca/census-recensement/2011/geo/bound-limit/bound-limit-2011-eng.cfm


