---
title: "Metro Van - rent analysis"
output: html_notebook
---

Author: Martin Monkman
Date: 2016-12-06

---

This file reads rent data (advertisements) from the Metro Vancouver region.




```{r}

library(dplyr)
library(readr)

# turn off scientific notation (where very large numbers write with e instead) in R
options(scipen = 999)

# set working director
setwd("C:/@data/GitHub/MetroVan_rent")


# define column names
columnnames <- c("reg", "latitude", "longitude", "rent", "bedrooms", "bathrooms", "sq_ft")

# ---------------

# read files, add column names
data1 <- read_csv("data/cl_2016_4_3.csv", col_names = FALSE)
colnames(data1) <- columnnames

data2 <- read_csv("data/cl_2016_4_4.csv", col_names = FALSE)
colnames(data2) <- columnnames

# bind the two data frames
# https://rdrr.io/cran/dplyr/man/bind.html
f01_data <- unique(bind_rows(data1, data2))

# ----------------
# a bunch of test code

#data2 <- read_csv("data/cl_2016_4_6.csv", col_names = FALSE)
#colnames(data2) <- columnnames
## bind the two data frames
## https://rdrr.io/cran/dplyr/man/bind.html
#f01_data <- unique(bind_rows(f01_data, data2))
#
# proto-loop
#i <- 11
#
#filename <- paste("data/cl_2016_4_", i, ".csv", sep="")
#datai <- read_csv(filename, col_names = FALSE)
#colnames(datai) <- columnnames
#
#head(datai)
#  
#datai$sq_ft <- gsub("open house dates", NA, datai$sq_ft)
#datai$sq_ft <- as.numeric(datai$sq_ft)
#
#  f01_data <- unique(bind_rows(f01_data, datai))


# ----- 
# loop! 
# (shout out to @ateucher  ¯\_(ツ)_/¯  )
# 
# IMPORTANT NOTE:
# this uses the April data
# the files 2016_4_8.csv and 2016_4_26.csv were empty;
# to make the loop work, a file with a set of comma separators was created

for (i in 5:30){
  filename <- paste("data/cl_2016_4_", i, ".csv", sep="")
  datai <- read_csv(filename, col_names = FALSE)
  colnames(datai) <- columnnames
  datai$sq_ft <- gsub("open house dates", NA, datai$sq_ft)
  datai$sq_ft <- as.numeric(datai$sq_ft)

  f01_data <- unique(bind_rows(f01_data, datai))
}


# add numeric rent field
f01_data$rent_num <- as.numeric(gsub('\\$','', f01_data$rent))


# filter out records that are not complete (i.e. NA values in any field, no geo-reference)
f01_data <- f01_data[complete.cases(f01_data), ]
f01_data <- f01_data %>%
  filter(latitude != 0)


write_csv(f01_data, "data/f01_data.csv")


```




#### Filtration

```{r}


# filter so that bedrooms = 1 or 2, bathrooms = 1
f03_data <- f02_data %>%
  filter(bedrooms != 0)

f03_data_1.2br <- f03_data %>%
  filter(bedrooms <= 2 & bathrooms == 1)


```

## Spatial data (mapping)


### Read in the shape file

For this exercise, we'll use the census tract shape files for the 2011 Census of Canada.


**References**

http://www.studytrails.com/r/r-and-gis-working-with-shapefiles/



```{r}
#
library(maptools)

library(sp)
library(rgeos)
#library(rgdal)

# from http://www.studytrails.com/r/r-and-gis-working-with-shapefiles/

crswgs84=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# read shapefile
BC_censustracts <- readShapePoly("CTshape/CT_2011.shp",proj4string=crswgs84,verbose=TRUE) 
 
# examine shapefile
class(BC_censustracts)
str(BC_censustracts@data)

# plot shapes
plot(BC_censustracts)


```


Zoom in on Metro Vancouver

Simple: just define the x and y limits using longitude (as x) and latitude (as y)

```{r}

# plot shapes
plot(BC_censustracts, 
     xlim=c(-123.3,-122.7), ylim=c(48.97,49.4))

```


Zoom in on the area covered by Skytrain routes

Simple: just define the x and y limits using longitude (as x) and latitude (as y)

```{r}

# plot shapes
plot(BC_censustracts, 
     xlim=c(-123.28,-122.79), ylim=c(49.14,49.31))

```

### Spatial join

The next step is to use the census tract shape file to tell us the census tract the rental unit (in the data file f03_data_1.2br).

This requires a spatial join, perhaps using the `over` command in the package sp.

reference: http://www.nickeubank.com/wp-content/uploads/2015/10/RGIS2_MergingSpatialData_part1_Joins.html

- using `point.in.polygon()` in `sp` package


Further details:
CRS - Coordinate Reference System (CRS)


---

#### futher reading

http://benrifkind.github.io/Mapping-Stats-Canada-Data-with-R-Part-3-of-3/


https://csde.washington.edu/services/gis/workshops/SPATIALR.shtml 

http://gis.stackexchange.com/questions/19064/how-to-open-a-shapefile-in-r


http://gis.stackexchange.com/questions/137621/join-spatial-point-data-to-polygons-in-r

https://cran.r-project.org/web/packages/sp/index.html 

http://stackoverflow.com/questions/24174042/matching-georeferenced-data-with-shape-file-in-r


**Statistics Canada shapefiles**

https://www12.statcan.gc.ca/census-recensement/2011/geo/bound-limit/bound-limit-2011-eng.cfm


